Post-Disturbance Tree Community Trajectories in a Neotropical Forest

### 2018-06-23 WholePlotTrajectories.tex
Update WholePlotTrajectories.tex

In the inventories from 1989 (5 years after disturbance) to 2015 (31
years after disturbance), 828388 trees and 591 botanical species
pertaining to 223 genus and 64 botanical families were recorded.
Communities taxonomic diversity trajectories were examined through the
Richness, Shannon and Simpson diversities at genus level, in relation to
the 1989 inventories (5 years after disturbance) (See annexe I). For
undisturbed plots the Richness, Shannon and Simpson diversities remained
stable over the 30 years monitored. In disturbed communities the
taxonomic richness increased after low disturbance intensity, reaching a
maximum gain of 14 botanical genera (plot 3 from treatment 2) while it
followed unimodal trajectories after intense disturbance, decreasing for
ten years before recovering pre-disturbance values. In all disturbed
plots the taxonomic evenness (Shannon and Simpson diversities)
increased, following unimodal trajectories with a maximum, reached after
around 20 years, positively correlated to the disturbance treatment
(\(\rho_{spearman}^{Shannon}=0.86\), and
\(\rho_{spearman}^{Simpson}=0.89\)). Return towards initial evenness
values was beginning after 30 years except for two T3 plots (plots 8 and
12) which evenness still increased, suggesting similar but delayed
trajectories \ref{fig:IDHplot}.

@@

The trajectories of taxonomic and functional composition were visualized
in a two dimensional ordination space mapping the successive inventories
according to their flora and corresponding traits. Classifications were
performed using either abundance-based Bray-Curtis (Figure
\ref{fig:NMDSplans}) or incidence-based Jaccard dissimilarity, both
giving similar results only analysis using Bray-Curtis dissimilarity are
discussed.

While both taxonomic and functional composition remained stable in
undistrubed communities, they followed consistent trajectories over time
after disturbance which revealed significant compositional changes.
According to the mapping of functional traits (see appendix I) these
compositional changes corresponded to shifts towards species with more
acquisitive functional strategies, from communities with high average WD
to high average SLA and chlorophyll content. For disturbed communities
the distance of successive inventories to the 1989 reference inventory
followed unimodal trajectories translating cyclic compositional changes
with a recovery of the initial composition (Figure \ref{fig:NMDSplans}).
The maximum dissimilarity with the initial state was positively
correlated to the disturbance treatment for both taxonomic and
functional composition (\(\rho_{spearman}^{taxonomic}=0.91\) and
\(\rho_{spearman}^{functional}=0.96\) respectively) and the time at
maximum was reached around 26 years after disturbance for taxonomic
composition and 22 years for functional composition.

@@

The changes observed in plots functional composition went hand to hand
with consistent trajectories of the 8 functional and life history traits
visualized with the trajectories of community weighted means (CWM) of
leaves economics (leaves thickness, chlorophyll content, toughness and
specific area), wood economics (wood specific gravity, bark thickness),
and life history traits (seed mass and maximum adult height) (Figure
\ref{fig:CWM}).

@@

Communities functional redundancy was measured as the sum within
communities of species functional overlap weighted by their abundance.
The functional redundancy was measured in the functional space defined
by 7 leaf, stem, and maximum height traits (see appendix I for PCA
details). Communities functional redundancy remained stable in control
plots but after disturbance the redundancy trajectories were quite
variable (See appendix I) and apparently independently of the initial
disturbance. Globally after most intense disturbance (plots T2 and T3)
communities redundancy decreased at first place before increasing to
edge, recover or exceed the initial value.

>>>>

From 1989 (2-3 years after disturbance) to 2015 (28-29
years after disturbance), 828388 individual trees and 591 botanical species
pertaining to 223 genus and 64 botanical families were recorded.
For undisturbed plots Richness, Shannon and Simpson diversities remained
stable over the 30 years monitored. In disturbed communities the
taxonomic richness increased after low disturbance intensity, reaching a
maximum gain of 14 botanical genera (plot 3 from treatment 2) while it
followed unimodal trajectories after intense disturbance, decreasing for
ten years before recovering to pre-disturbance values. In all disturbed
plots the taxonomic evenness (Shannon and Simpson diversities)
increased, following unimodal trajectories with a maximum, reached after
around 20 years, positively correlated to the disturbance treatment
(\(\rho_{spearman}^{Shannon}=0.86\), and
\(\rho_{spearman}^{Simpson}=0.89\)). Return towards initial evenness
values was beginning after 30 years except for two T3 plots (plots 8 and
12) which evenness still increased, suggesting similar but delayed
trajectories \ref{fig:IDHplot}.

@@

The trajectories of taxonomic and functional composition were visualized
in a two dimensional ordination space mapping the successive inventories
according to their flora and corresponding traits. Classifications were
performed using either abundance-based Bray-Curtis (Figure
\ref{fig:NMDSplans}) or incidence-based Jaccard dissimilarity, both
giving similar results only analysis using Bray-Curtis dissimilarity are
presented.

While both taxonomic and functional composition remained stable in
undisturbed communities, they followed consistent trajectories over time
after disturbance which revealed significant compositional changes.
According to the mapping of functional traits (see appendix I) these
compositional changes corresponded to shifts towards species with more
acquisitive functional strategies, from communities with high average WD
to high average SLA and chlorophyll content. For disturbed communities
the distance of successive inventories to the 1989 reference inventory
followed unimodal trajectories translating cyclic compositional changes
with a recovery of the initial composition (Figure \ref{fig:NMDSplans}).
The maximum dissimilarity with the initial state was positively
correlated to the disturbance treatment for both taxonomic and
functional composition (\(\rho_{spearman}^{taxonomic}=0.91\) and
\(\rho_{spearman}^{functional}=0.96\) respectively) and the time at
maximum was reached around 26 years after disturbance for taxonomic
composition and 22 years for functional composition.

@@

The changes observed in plots functional composition went hand to hand
with consistent trajectories of the 8 functional traits
visualized with the trajectories of community weighted means (CWM) of
leaves economics (leaves thickness, chlorophyll content, toughness and
specific area), wood economics (wood specific gravity, bark thickness),
and life history traits (seed mass and maximum adult height) (Figure
\ref{fig:CWM}).

@@

Communities functional redundancy remained stable in control
plots but after disturbance the redundancy trajectories were quite
variable (See appendix I) and apparently independently of the initial
disturbance. Globally after most intense disturbance (plots T2 and T3)
communities redundancy decreased at first place before increasing to
edge, recover or exceed the initial value.
